---
title: "Class11: structural Bioinformatics"
author: "Negin Silani"
date: "5/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##The PDBB database

The [PDB](http://www.rcsb.org/) is the main repository for biomolecular structural data.

Here we examine the contents of the PDB:

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein? Aim to have a rendered GitHub document
with working code that yields your answers

```{r}
db<- read.csv("Data Export Summary.csv", row.names=1)
head(db)
```

How many are X-Ray, etc..
```{r}
(db$Total/sum(db$Total)) * 100
```

What percent are proteins..

```{r}
(sum(db$Proteins)/sum(db$Total)) * 100
```

we could also try the datapasta package and copy from website and "Addins" > "Paste as data.frame..

```{r}
library(datapasta)
        
tmp<- data.frame(stringsAsFactors=FALSE,
                     Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                                             "Multi Method", "Total"),
                                Proteins = c(126880, 11062, 2277, 256, 129, 140604),
                           Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
                      ProteinComplex = c(6547, 259, 800, 6, 2, 7614),
                                   Other = c(8, 8, 0, 13, 1, 30),
                                   Total = c(135447, 12608, 3108, 279, 137, 151579)
                  )
```

Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

There are 1157 as of 2019-05-07 see:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=880B2268


##Section 3 Using Bio3D

```{r}
library(bio3d)

pdb<- read.pdb("1hsg.pdb")
pdb
```

```{r}
atom.select(pdb, "protein", value= TRUE)
```


Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues?

amino acid residues: 1686
two non protein residues: HOH (127), MK1 (1)

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

# Print a subset of $atom data for the first two atoms

```{r}
aa321(pdb$atom$resid)
```


```{r}
atom.select(pdb, "protein")
```

```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```

# Note that individual $atom records can also be accessed like this

```{r}
pdb$atom$elety[1:2]
```

# Which allows us to do the following

```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab=“B-factor”)
```

Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a
useful summery of any R object.

# Print a summary of the coordinate data in $xyz

```{r}
pdb$xyz
```

Atom selection is done via the function **atom.select()**

```{r}
prot.pdb<- atom.select(pdb, "ligand", value= TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

```{r}
#pdb$atom[inds$atom, ]
```

```{r}
lig.pdb<- atom.select(pdb, "ligand", value= TRUE)
write.pdb(lig.pdb, file="1hsg_protein.pdb")
```

```{r}
aa <- get.seq("1ake_A")
```


```{r}
aa
```

# Blast or hmmer search 
```{r}
b <- blast.pdb(aa)
```

```{r}

```

