---
title: "Class7. R functions and packages"
author: "Negin Silani"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions revisited

we will source a file from online with our functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function


#Find missing NA values in two vectors

start with a simple example of the larger problem I am trying to solve

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```

take the sum() to find out how many TRUE values we have and thus how many NAs we had in both x and y

```{r}
sum( is.na(x) & is.na(y))
```
Now I can make this into our first function..

```{r}
both_na<- function (x, y) {
  sum( is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA,3,NA,2,NA))
```



```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x,y2)
```

```{r}
y3<- c(1,NA,NA,NA,NA,NA,NA)
both_na(x,y3)
```


```{r}
3 == 3
```

```{r}
3<2
```

```{r}
3!= 2
```

```{r}
length(x)
```

```{r}
length(y2)
```


```{r}
#both_na2(x, y2)
```

```{r}
which( c(FALSE,FALSE,TRUE,FALSE,TRUE) )


#which( is.na( c(1,2, NA,4)) )
```



```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na3(x, y)
```

##insert function

```{r}
df1
```

```{r}
df2
```

```{r}
# Start with a simple version of the problem
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
 exp=c(2,1,1),
 stringsAsFactors=FALSE)
df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
 exp=c(-2, NA, 1, 2),
 stringsAsFactors=FALSE)
```

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect(x, y)
```

```{r}
(x %in% y)
```

```{r}
x[x %in%y]
```

Use the Rstudio shortcut ' CODE> EXTRACT FUNCTION' to turn our snippet into a working function
```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[y %in% x])
}
```


```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1,df2, by="IDs")
```

##Grade function

```{r}
x1<- c(100, 100, 100, 100, 100, 100, 100, 90)
x2<- c(100, 90, 90, 90, 90, 90, 97, 80)
```

Find average grade dropping the worst homework score
```{r}
min(x1, na.rm=FALSE)
(sum(x1)- min(x1))/7
```

```{r}
#code (x1) grade<- function(x1) {
 # min(x1, na.rm=FALSE)
  #(sum(x1)- min(x1))/7
#}
```

```{r}
min(x2, na.rm= FALSE)
```

```{r}
(sum(x2) - min(x2))/7
```

```{r}
#code (x2) grade <- function(x2) {
  #(sum(x2) - min(x2))/7
#}
```

